---
title: "Economic Underwriting Aid"
output: 
  flexdashboard::flex_dashboard:
    theme: 
      version: 4
runtime: shiny
---

```{r global}
# Load required packages
library(shiny)

library(plotly)

library(DT)

library(leaflet)

library(leaflet.providers)

library(leaflet.extras)

library(sf)

library(dplyr)

library(RColorBrewer)

library(shinydashboard)

Entire_US_1yr_ACS <- readRDS("EntireUS_5yr_ACS.RDS")
col_indices = as.character(6:ncol(Entire_US_1yr_ACS))
```


# Inputs

Here is where you can input the specific geographics and demographics you would like to see.  You can select multiple states, counties, and layers depending on what you are interested in.  The numbers of the layers correspond to the variables displayed in the variable table

```{r}
selectInput("state", "Select State:", choices = unique(Entire_US_1yr_ACS$State), multiple = TRUE)
selectInput("county", "Select County:", choices = NULL, multiple = TRUE)
selectInput("column1", "Select Layer 1:", choices = col_indices)
selectInput("column2", "Select Layer 2:", choices = col_indices)
selectInput("column3", "Select Layer 3:", choices = col_indices)
downloadButton('downloadData', 'Download All Data')
actionButton("render_map", "Render Map")
fileInput("fileUpload", "Upload Modified CSV")
```

# Variable Table

Each variable corresponds to a number, only numerical variables can be selected.  For example, if you want to see the number of house holds who are making 75-99k, 100-149k, and 150-199k then you would set the Layer inputs to 18, 19, 20

```{r}
DTOutput("table2")
```

# Map

```{r}
leafletOutput("map1",width = "100%")
```

# Graph!

```{r}
#plotlyOutput("plotlyBarChart",width = "100%")
```


# Map Data

```{r}
DTOutput("table")
```


```{r}
    # Update County Dropdown based on selected States
    observe({
      filtered_counties <- Entire_US_1yr_ACS %>% 
        filter(State %in% input$state) %>% 
        pull(County) %>% 
        unique()
      updateSelectInput(session, "county", choices = filtered_counties)
    })
    
    # Initialize a reactive value to hold the filtered data
    filtered_data_by_state_county <- reactiveVal()
    
    # Update the filtered data only when the new button is clicked
    observeEvent(input$render_map, {
      filtered_data_by_state_county(Entire_US_1yr_ACS %>% 
                                      filter(State %in% input$state, 
                                             County %in% input$county))
    })
    # Render Leaflet Map
    output$map1 <- renderLeaflet({
      # Filter data based on input
      filtered_data <- filtered_data_by_state_county()
      
      # Generate map
      pal1a <- colorNumeric("Blues", domain = filtered_data[[as.integer(input$column1)]])
      pal1b <- colorNumeric("Reds", domain = filtered_data[[as.integer(input$column2)]])
      pal1c <- colorNumeric("Greens", domain = filtered_data[[as.integer(input$column3)]])
      pal2 <- colorFactor("Set1", domain = filtered_data$County)
      
      leaflet(filtered_data) %>%
        addProviderTiles("OpenStreetMap.Mapnik") %>%
        addPolygons(
          fillColor = ~pal1a(filtered_data[[as.integer(input$column1)]]),
          weight = 0.5,
          opacity = 0.4,
          color = "black",
          fillOpacity = 0.7,
          group = ~paste(input$column1),
          popup = ~paste(input$column1, ":",filtered_data[[as.integer(input$column1)]], "<br>", County, "<br>", State, "<br>", Tract)
        ) %>%
        addPolygons(
          fillColor = ~pal1b(filtered_data[[as.integer(input$column2)]]),
          weight = 0.5,
          opacity = 0.4,
          color = "black",
          fillOpacity = 0.7,
          group = ~paste(input$column2),
          popup = ~paste(input$column2, ":", filtered_data[[as.integer(input$column2)]], "<br>", County, "<br>", State, "<br>", Tract)
        ) %>%
        addPolygons(
          fillColor = ~pal1c(filtered_data[[as.integer(input$column3)]]),
          weight = 0.5,
          opacity = 0.4,
          color = "black",
          fillOpacity = 0.7,
          group =  ~paste(input$column3),
          popup = ~paste(input$column3, ":", filtered_data[[as.integer(input$column3)]], "<br>", County, "<br>", State, "<br>", Tract)
        ) %>%
        addPolygons(
          fillColor = ~pal2(County),
          weight = 0.5,
          opacity = 0.4,
          color = "black",
          fillOpacity = 0.7,
          group = "County Layer",
          popup = ~paste(County, "<br>", State, "<br>", Tract)
        ) %>%
        addLayersControl(
          overlayGroups = c(~paste(input$column1),~paste(input$column2),~paste(input$column3), "County Layer"),
          options = layersControlOptions(collapsed = FALSE)
        )
    })
    # Render Data Table
    output$table <- renderDT({
      datatable(data.frame(filtered_data_by_state_county()), filter = 'top', extensions = c('Buttons'), editable = TRUE,
                options = list(
                  dom = 'Bfrtip',
                  buttons = list(
                    list(
                      extend = 'copy',
                      exportOptions = list(
                        columns = ':visible',
                        modifier = list(page = 'current')
                      )
                    ),
                    list(
                      extend = 'csv',
                      exportOptions = list(
                        columns = ':visible'
                      )
                    ),
                    list(
                      extend = 'excel',
                      exportOptions = list(
                        columns = ':visible'
                      )
                    )
                  ),
                  lengthMenu = c(5, 10, 15, 20),
                  pageLength = 50,
                  scrollX = TRUE
                )
      )
    })
    
    # Render Data Table
    output$table2 <- renderDT({
      datatable(data.frame(data.frame(`Layers` = colnames(filtered_data_by_state_county()))), filter = 'top', extensions = c('Buttons'), editable = TRUE,
                options = list(
                  dom = 'Bfrtip',
                  buttons = list(
                    list(
                      extend = 'copy',
                      exportOptions = list(
                        columns = ':visible',
                        modifier = list(page = 'current')
                      )
                    ),
                    list(
                      extend = 'csv',
                      exportOptions = list(
                        columns = ':visible'
                      )
                    ),
                    list(
                      extend = 'excel',
                      exportOptions = list(
                        columns = ':visible'
                      )
                    )
                  ),
                  lengthMenu = c(5, 10, 15, 20),
                  pageLength = 50,
                  scrollX = TRUE
                )
      )
    })
    
    output$downloadData <- downloadHandler(
      filename = function() { paste('data-', Sys.Date(), '.csv') },
      content = function(file) {
        write.csv(as.data.frame(filtered_data_by_state_county()) %>% select(-geometry), file)
      }
    )
```


```{r}
# # Render Plotly bar chart for column1 vs. Census Tract
# output$plotlyBarChart <- renderPlotly({
#   # Access the reactive filtered data
#   filtered_data <- filtered_data_by_state_county()
#   
#   # Check if data is available
#   if (is.null(filtered_data)) {
#     return(NULL)
#   }
#   
#   # Access the specific column (column1) from user input
#   column_data <- filtered_data[[as.integer(input$column1)]]
#   
#   # Create Plotly bar chart
#   plot <- plot_ly(filtered_data, x = ~GEOID, y = filtered_data[[as.integer(input$column1)]], type = 'bar') %>%
#     layout(title = paste("Bar Chart of", input$column1, "by Census Tract"),
#            xaxis = list(title = "Census Tract"),
#            yaxis = list(title = input$column1))
#   
#   return(plot)
# })
```

